---
title: "Lab 2 Analysis Report - High School Education and
its Effects on GPA"
author: "Robert Cocker"
date: "2/8/2025"
params:
    college: "The University of Texas at Arlington"
    college_class: "DATA 3442 - Statistical Methods in Data Science 2"
output: pdf_document
---
```{r setup, include=FALSE}
# R Markdown pre-checks (R Markdown file processing)
# Load necessary libraries
library(rmarkdown)
library(psych)
library(pastecs)
library(corrplot)
library(GGally)
library(ggplot2)

# Set global chunk options
knitr::opts_chunk$set(echo = TRUE)
```
---
**College:** `r params$college`  
**Class:** `r params$college_class`

## Lab Study Introduction

### This report analyzes the relationship of high school education, SAT scores, and its effects in predicting cumulative GPA.
#### Run code blocks for inference and analyis of this study.

## Statistical Regression Methods: Data Summary
### Summary of Descriptive Statistics
```{r}
summary(gpa_data)
describe(gpa_data)
stat.desc(gpa_data)
```

### Correlation Matrix
```{r}
library(corrplot)
corrplot(cor(gpa_data[, -1]), method = "color", tl.cex = 0.8, order = "hclust")
```

### Scatterplot Matrix
```{r}
ggpairs(gpa_data,lower = list(continuous = wrap("smooth", colour = "blue")))
```

## Regression Models
### High School Model
```{r}
summary(hs_model)
```

### SAT Model
```{r}
summary(sat_model)
```

### Full Model
```{r}
summary(full_model)
```

## Model Comparisons
### ANOVA: High School vs Full Model
```{r}
anova(hs_model, full_model)
```

### ANOVA: SAT vs Full Model
```{r}
anova(sat_model, full_model)
```

### Stepwise Regression
```{r}
summary(stepwise_model)
```

### Final Regression Model (most robust)
```{r}
summary(final_model)
```

## Model Diagnostics
```{r}
par(mfrow = c(2,2))
plot(full_model)
```

## Regression Plots
```{r}
# GPA vs HSM Model
plot_hsm <- ggplot(gpa_data, aes(x = hsm, y = gpa)) + 
  geom_point() + 
  geom_smooth(method = "lm", col = "blue") + 
  ggtitle("GPA vs High School Math (HSM)")
print(plot_hsm)

# GPA vs HSS Model
plot_hss <- ggplot(gpa_data, aes(x = hss, y = gpa)) + 
  geom_point() + 
  geom_smooth(method = "lm", col = "blue") + 
  ggtitle("GPA vs High School Science (HSS)")
print(plot_hss)

# GPA vs HSE Model
plot_hse <- ggplot(gpa_data, aes(x = hse, y = gpa)) + 
  geom_point() + 
  geom_smooth(method = "lm", col = "blue") + 
  ggtitle("GPA vs High School English (HSE)")
print(plot_hse)

# GPA vs SATM Model
plot_satm <- ggplot(gpa_data, aes(x = satm, y = gpa)) + 
  geom_point() + 
  geom_smooth(method = "lm", col = "red") + 
  ggtitle("GPA vs SAT Math (SATM)")
print(plot_satm)

# GPA vs SATV Model
plot_satv <- ggplot(gpa_data, aes(x = satv, y = gpa)) + 
  geom_point() + 
  geom_smooth(method = "lm", col = "red") + 
  ggtitle("GPA vs SAT Verbal (SATV)")
print(plot_satv)

# Close the PDF device
dev.off()
```

# Conclusion of study
The final regression model summary is provided. See `analysis_results` and pdf files for more information. Leveraging stepwise regression modeling, we confirm that High School Math (HSM) has a statistically significant effect on cumulative GPA (p < 0.001), while High School English (HSE) may have a small but unreliable effect (p = 0.1637). Other predictors of High School Science (HSS) and SAT scores (SATM, SATV), do not significantly contribute to GPA prediction. These results suggest that strong math performance in high school is a key driver of overall academic success, while standardized test scores are less relevant in explaining GPA variations. Our final model has an R^2 of ~.202 which explains about 20.2% of the model from its data. This contributes to the study and body of research for high school education stateside and abroad.  

This conclusion is based on robust statistical evidence from multiple regression models, hypothesis testing, and model selection methods, reinforcing its validity and reliability.

### Render R Markdown to documentation file format(s)
To render the lab report the .Rmd R Markdown file must be rendered via R console. File output format for documentation can be changed.See link for more information: https://pkgs.rstudio.com/rmarkdown/reference/render.html

### Final Thoughts
R has Linux functionality and fragmentation issues in its stack. This lab was completed on RStudio GUI due to technical difficulties on a another machine on Linux in Bash. With that said, R seems to be a very robust programming language.

## References

Course and Online sources:

- Lecture, lab, and course materials
- ChatGPT
- Google

R Documentation:

- [R Project Documentation](https://www.r-project.org/other-docs.html)  
- [RDocumentation.org](https://www.rdocumentation.org/)